---
title: "Lecture 2"
output: html_notebook
---

```{r}
library(tidyverse)
library(nycflights13)
```
# LOADING DATA

We will load the flights of 3 different airport in 2013 from NYC

```{r}
flights <- nycflights13::flights
head(flights)
```
## FILTER

```{r}
filter(flights, month==3, day>=15)
```
## ARRANGE/SORTING

```{r}
arrange(flights,month, desc(dep_delay))
```

```{r}
names(flights)
```
## SELECT

```{r}
select(flights, tailnum, dep_time)
```

## ADD NEW DATA by deriving information
### MUTATE
We want to add a column with the average speed in minutes
```{r}
mutate(flights, avg_speed = distance / air_time * 60)
```

### MUTATE with IFELSE

```{r}
my_flights <- select(flights, dep_delay)
mutate(my_flights, if_else(dep_delay > 0,
                           "delayed", # if true
                           "on_time", # if else
       ))
```

# SELECT COUNT * FROM ... GROUP BY ...
This is an sql analogy.

This is a first instance of aggregation.

```{r}
count(flights, month)
flights_by_month <- group_by(flights, month)
```

# SUMMARIZE

```{r}
summarize(flights_by_month, avg_delay = mean(dep_delay, na.rm=TRUE))
```

```{r}
flights |> group_by(month) |> summarise(avg_delay=mean(dep_delay, na.rm=TRUE))
```

# Exercise 1

Compute the fraction of delay flights per month.

```{r}
flights |>
  group_by(month) |>
  select(month,dep_delay) |>
  mutate(is_delayed = if_else(dep_delay>0,
                 1, #if true
                 0, #if false
                 )) |>
  summarise(fraction_delay_flights=mean(is_delayed, na.rm=TRUE))
```

# JOINS

```{r}
my_flights <- select(flights, year, origin, dest, tailnum)
```
```{r}
planes <- nycflights13::planes
my_planes <- select(planes, tailnum, year, manufacturer, model)
```

```{r}
inner_join(my_flights, my_planes, by=c("tailnum"), suffix=c("_flight","_manufacture"))
```


