---
title: "Lecture 6 - Color"
output: html_notebook
---

# Loading data and preparing workspace
```{r}
library("colorspace")
library("viridis")
library("RColorBrewer")
library("ggthemes")
```

# 
```{r}
pal_viridis <- viridis::cividis(6)
pal_set2 <- RColorBrewer::brewer.pal(6, "Set2")
# Most used categorical palette
pal_tableau <- ggthemes::tableau_color_pal()(6)
pal_viridis
```
```{r}
colorspace::swatchplot(
  "Viridis" = pal_viridis,
  "Brewer" = pal_set2,
  "Tableau" = pal_tableau
  )
```
## Assessing quantitatively palettes
```{r}
colorspace::specplot(pal_tableau)
```

## Assessing quality of palettes for color blind people
```{r}
pal_deu_tableau <- colorspace::deutan(pal_tableau)
```
```{r}
pal_deu_tableau |>
  colorspace::specplot()
```
```{r}
colorspace::swatchplot(
  "no cvd"=pal_tableau,
  "cvd"=pal_deu_tableau
)
```
###Protan blindness
```{r}
pal_viridis |>
  protan() |>
  specplot()
```

### Tritan blindness
```{r}
pal_viridis |>
  tritan() |>
  specplot()
```
## Grey-White 
```{r}
pal_viridis |>
  desaturate(1) |>
  specplot()
```

```{r}
pal_tableau |>
  desaturate(1) |>
  specplot()
```
```{r}
rainbow(6) |>
  specplot()
```

# Quantitative assessment of contrast

Contrast ratio: luminance of the color / luminance of the background

What are good colors? Around 3 is lower bound

```{r}
colorspace::contrast_ratio(pal_viridis, col2="grey30", plot=TRUE)
```
# How can we fix tableau for blind people? 
Therefore cathegorical data for blind people safe.
```{r}
# Okabe and Ito - Japanese professors inventors
palette.colors(6) |>
  desaturate(1) |>
  specplot()
```

# Now we play
```{r}
pal_random <- c("#ff6f59", "#254441", "#43aa8b", "#b2b09b", "#ef3054")
pal_random |>
  specplot()
```
```{r}
#Then you can define the following function:
image_spec <- function(img_path, palette) {
  p_image <- cowplot::ggdraw() +
    cowplot::draw_image(img_path)

  p_palette <- tibble(c = palette) |>
    ggplot(aes(x=palette, y=0, fill=palette)) +
    geom_tile(color="white", linewidth=4) +
    coord_fixed() +
    scale_fill_identity() +
    theme_void()

  plot_grid(p_palette, p_image, ncol=1)
}
```


```{r}
pal_dune <- c("#D96A6A","#8C4660","#592550","#F8AEA1","#260101")
img_path <- "data/download.jpg"
image_spec(img_path, pal_dune)
```

```{r}
pal_dune |> 
  desaturate(1) |>
  specplot()
```
```{r}
pal_burano <- c("#B6DBF2","#049DBF","#055902","#F2B705","#F2CDAC")
img_path <- "data/images.jpg"
image_spec(img_path, pal_burano)
```
# Another complicate palette
```{r}
base1 <- "#C91024"
base2 <- "#1E8CE3"
saturations <- c(0.0, 0.5, 1.0)
lightenesses <- c(0.0, 0.3, 0.8)
```
```{r}
base1 |>
  desaturate(0.5) |>
  lighten(0.3) |>
  swatchplot()
```
```{r}
color_base_tibble <- function(base, saturation, lightness){
  tibble(c=base,
         sat=saturation,
         light=lightness) |>
    transmute(
      c=lighten(desaturate(c,sat),light),
      c=fct_reorder(c, row_number())
    )
}
```
```{r}
cols1 <- base1 |>
  color_base_tibble(saturations, lightenesses) |>
  rename(C1=c)
base2 |>
  color_base_tibble(saturations, lightenesses) |>
  rename(C2=c)
```

```{r}
blend_colors <- function(C1,C2){
  hex(RGB(coords(hex2RGB(C1)) * coords(hex2RGB(c2)) ))
}
```
```{r}

```



